---
title: "Flux analysis"
date: "10/11/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r libraries}
library(dplyr)
library(tidyr)
library(ggplot2)
library(naniar)
library(ggpubr)
library(patchwork)
library(PerformanceAnalytics)
```

**Meteorological data:** 
(UFORA) You only need the data corresponding to your study site.

- Time of the measurement (YEAR, MONTH, and DAY)
- Daily average temperature T (°C)
- Daily average night temperature, T_NIGHT (°C)
- Daily average day temperature, T_DAY (°C)
- Incoming shortwave radiation, SWDOWN (W/m²)
- Incoming longwave radiation, LWDOWN (W/m²)
- Vapor pressure deficit, VPD (hPa)
- Rainfall, RAIN (mm)
- Windspeed, WIND (m/s)
- Net radiation, RN (W/m²)
- Soil temperature, TSOIL (°C)
- Soil heat flux, GF (W/m²)
- Latent heat flux QLE (W/m²)
- Sensible heat flux QH (W/m²)
- Net Ecosystem Exchange, NEE (gC/m²/d)
- Gross primary production, GPP (gC/m²/d)
- Ecosystem respiration, Reco (gC/m²/d)

**Additional helpful Metadata:**

- World map of Köppen-Geiger climate classification (PDF)
- Site information: Find from [fluxnet sites list](https://fluxnet.org/sites/site-list-and-pages) or just google the name of your site


### Step 1: Explore your dataset 
Data cleanup, units conversion; typical ranges, annual sums and averages  

- Check and filter missing (NA) or impossible (e.g. very larges, -9999) values
- Convert variables into easy interpretable units (land fluxes into kgC /m²/year or tC /ha/year)
- Calculate the minimal and maximal value for each variable, as a quick check
- Calculate annual sum of precipitation, and the annual mean for each other variable. 

```{r Loading your data}
# Loading data
Fluxname <- "ZM-Mon_daily"
data.file <- file.path("./",paste0(Fluxname,".csv"))
data <- read.csv(data.file,header = TRUE,sep = ";")

```

```{r cleaning data}
# cleaning the data
data <- data%>%
  replace_with_na_if(.predicate = is.double,condition = ~.x < -1000)


```

```{r Unit conversion}
# Editing the units
data <- data %>%
  mutate(
    NEE = NEE * 10^-3 * 365,
    RECO = RECO * 10^-3 * 365,
    GPP = GPP * 10^-3 * 365
  )
min <- data.frame(as.list(apply(data, 2, min, na.rm=TRUE)))
max <- data.frame(as.list(apply(data, 2, max, na.rm=TRUE)))
summarie <- rbind(min, max)

precip_cumsum <- data %>%
  group_by(YEAR)%>%
  summarise(sum=sum(RAIN))

data$Date <- as.Date(paste(data$YEAR, data$MONTH, data$DAY, sep = "-"))
```


### Step 2: Mean annual cycles
Seasonal cycle, patterns, extreme events 


**Plot 1**: Produce mean seasonal cycle graphs (= you compile your data to obtain one average year) for each meteorological variable (in a single plot)


- Check the correspondence of these plots: do patterns concur or differ? Why? Try to explain the causal relationship. 
- **Plot 2**: Compare your average year with every single year of data. Are there any extreme events in your data? Document it.
- **Plot 3**: Using bar plots, plot the average precipitation yearly cycle
```{r plot 1}
data$DayOfYear <- as.numeric(format(data$Date, format = "%j"))

averages_per_day_year <- data %>%
  group_by(DayOfYear) %>%
  summarize(across(c(T, T_NIGHT, T_DAY, SWDOWN, LWDOWN, VPD, RAIN, WIND, RN, TSOIL, GF, QLE, QH, NEE, RECO, GPP), mean, na.rm = TRUE))

averages_per_month <- data %>%
  group_by(MONTH) %>%
  summarize(across(c(T, T_NIGHT, T_DAY, SWDOWN, LWDOWN, VPD, RAIN, WIND, RN, TSOIL, GF, QLE, QH, NEE, RECO, GPP), mean, na.rm = TRUE))

meteorological_data <- averages_per_day_year[, c("DayOfYear", "T", "T_NIGHT", "T_DAY","SWDOWN","LWDOWN", "VPD", "RAIN", "WIND", "TSOIL")]%>%
  mutate(SWDOWN=SWDOWN/10,
         LWDOWN=LWDOWN/10)

meteorological_data_long <- pivot_longer(meteorological_data, cols = -DayOfYear, names_to = "Variable", values_to = "Value")

get_season <- function(day) {
  if(day >= 80 && day <= 172) return("herfst")
  else if(day > 172 && day <= 264) return("winter")
  else if(day > 264 && day <= 355) return("lente")
  else return("zomer")
}

meteorological_data_long <- meteorological_data_long %>%
  mutate(Season = sapply(DayOfYear, get_season))

#PLOT 1
ggplot(data=meteorological_data_long, aes(x = DayOfYear, 
                                     y = Value, 
                                     color = Variable)) +
  geom_line() +
geom_vline(xintercept=c(121,244))+
  labs(title = "Gemiddelde meteo variabelen per dag van het jaar",
       x = "Dag van het jaar",
       y = "gemiddelde waarde") +
  scale_color_manual(values=c('darkgreen','blue','green', 'red', 'orange', 'black', 'brown',
                              'purple','#FF00CC'), 
                     labels = c('LW down (daW/m²)', 'regen (mm/dag)', 'SW down (daW/m²)', 'T (°C)',
                                'dag T (°C)', 'nacht T (°C)', 'bodem T (°C)', 'VPD (hPa)',
                                'wind (m/s)'))+
  annotate("text", x=50, y=40, label = "nat seizoen")+
  annotate("text", x=180, y=40, label = "droog seizoen")+
  annotate("text", x=300, y=40, label = "nat seizoen")


#PLOT 2
extended_data <- merge(data, averages_per_day_year, by='DayOfYear')
extended_data$YEAR <- as.character(extended_data$YEAR)
colnames(extended_data) <- c('DayOfYear', 'jaar', 'maand', 'dag', 
                             'T (°C)', 'nacht T (°C)', 'dag T (°C)', 
                             'SW down (W/m²)', 'LW down (W/m²)', 'VPD (hPa)',
                             'regen (mm)', 'wind (m/s)','RN (W/m²)',
                             'bodem T (°C)', 'GF (W/m²)', 'QLE (W/m²)',
                             'QH (W/m²)','NEE (kgC/m²/jaar)', 
                             'RECO (kgC/m²/jaar)', 'GPP (kgC/m²/jaar)',
                             'Date','gem.T (°C)', 'gem.nacht T (°C)', 
                             'gem.dag T (°C)','gem.SW down (W/m²)',
                             'gem.LW down (W/m²)', 'gem.VPD (hPa)',
                             'gem.regen (mm)', 'gem.wind (m/s)',
                             'gem.RN (W/m²)','gem.bodem T (°C)',
                             'gem.GF (W/m²)', 'gem.QLE (W/m²)',
                             'gem.QH (W/m²)','gem.NEE (kgC/m²/jaar)', 
                             'gem.RECO (kgC/m²/jaar)', 'gem.GPP (kgC/m²/jaar)'
                             )

legende = c(rep(FALSE, 7),TRUE, rep(FALSE, 8))
for (i in 1:16){
  name <- paste0('plot_', i)
  assign(name, ggplot(extended_data, aes(x=.data[[colnames(extended_data)[21+i]]],
                                         y=.data[[colnames(extended_data)[4+i]]], 
                                         color=jaar))+
           geom_point(show.legend = legende[i])+
           geom_abline(slope=1, intercept=0)+
           labs(x = colnames(extended_data)[21+i],
                y = colnames(extended_data)[4+i]))
}


samen <- ggarrange(plot_1, plot_2, plot_3, plot_4,
                   plot_5, plot_6, plot_7, plot_8,
                   plot_9, plot_10, plot_11, plot_12,
                   plot_13, plot_14, plot_15, plot_16,
                   ncol = 4, nrow = 4, common.legend=TRUE)
samen

for (i in 1:16){
  name <- paste0('plot_', i)
  assign(name, ggplot(extended_data, aes(x=DayOfYear, 
                                         color=jaar))+
           geom_line(aes(y=.data[[colnames(extended_data)[21+i]]]), show.legend = legende[i])+
           geom_line(aes(y=.data[[colnames(extended_data)[4+i]]]), show.legend = legende[i])+
           labs(x = 'Dag van het Jaar',
                y = colnames(extended_data)[4+i]))
}


samen2 <- ggarrange(plot_1, plot_2, plot_3, plot_4,
                   plot_5, plot_6, plot_7, plot_8,
                   plot_9, plot_10, plot_11, plot_12,
                   plot_13, plot_14, plot_15, plot_16,
                   ncol = 4, nrow = 4, common.legend=TRUE)
samen2

#PLOT 3
average_rain_per_month <- data%>%
  group_by(YEAR,MONTH)%>%
  summarise(rain = sum(RAIN))%>%
  ungroup()%>%
  group_by(MONTH)%>%
  summarise(rain=mean(rain))
#average_rain_per_month <- average_rain_per_month[order(average_rain_per_month$MONTH),]
coeff <- 5
regen <- ggplot(average_rain_per_month, aes(x=MONTH, y=rain))+
  geom_histogram(stat='identity', fill='blue')+
  geom_line(data=averages_per_month, aes(x=MONTH, y=T*coeff), col = 'red')+
  geom_point(data=averages_per_month, aes(x=MONTH, y=T*coeff), col = 'red', fill='red', shape=23)+
  scale_x_discrete(limits=1:12,
                   name = 'Maand')+
  scale_y_continuous(name='gemiddelde regenval (mm)',
                   limits=c(0,150),
                   sec.axis = sec_axis( trans=~./coeff, name="Temperature (°C)"))
regen
```


### Step 3: Surface energy balance
Radiation balance, energy closure, Bowen ratio 

- **Plot 4**: Plot mean annual cycles for Rn, Qh, Qle, GF (W/m²), in one graph  and interpret the annual cycles.
- **Plot 5**: Scatterplot of energy balance (Qh+Qle vs. Rn-GF) (W/m²), select only 1 year of your dataset for this plot, fit a linear regression line (include equation + R²). Compare with 1:1 line.
- Analyze the closure of the energy balance. How much is the energy gap? What causes this? Did you expect such magnitude? Refer to literature in your answer.
- Calculate the Bowen ratio (= Qh/Qle).
```{r }
radiation_cycles <- data %>%
  group_by(DayOfYear)%>%
  summarise(mean_RN = mean(RN,na.rm=TRUE),
            mean_QH = mean(QH, na.rm=TRUE),
            mean_QLE = mean(QLE, na.rm=TRUE),
            mean_GF = mean(GF, na.rm=TRUE))
radiation_cycles <- pivot_longer(radiation_cycles, cols=contains('mean'),
                                 names_to = 'variables',
                                 values_to = 'values')
#PLOT 4
ggplot(data=radiation_cycles, aes(DayOfYear, values, colour=variables)) +
  geom_line()+
  scale_colour_discrete(labels=c('Ground Heat Flux', 'Sensible Heat Flux', 
                               'Latent Heat Flux', 'Net Radiation'),
                        name = 'variabelen')+
  labs(x='Dag van het Jaar',
       y= 'straling (W/m²)',
       title = 'stralingsverloop 2008')


#PLOT 5
P5_data <- data%>%
  filter(YEAR==2008)%>%
  mutate(QH_QLE = QH+QLE,
         RN_GF = RN-GF)
ggplot(P5_data, aes(x=RN_GF, y=QH_QLE))+
  geom_point()+
  geom_smooth(method = 'lm', se=FALSE, col = 'red')+
  stat_regline_equation(label.x = 0, label.y=170)+
  stat_cor(label.x = 0, label.y=180)+
  geom_abline(intercept=0, slope=1, linetype=2)+
  labs(x= 'Rn - G (W/m²)', y='Qle + Qh (W/m²)',
       title = 'Visualisatie Energieverlies 2008')

Bowen_ratio <- P5_data$QH/P5_data$QLE

P4_2_data <- cbind(P5_data$DayOfYear, Bowen_ratio)
colnames(P4_2_data) <- c('DayOfYear', 'BR')
P4_2_data <- as.data.frame(P4_2_data)
ggplot(P4_2_data, aes(x=DayOfYear, y=BR) )+
  geom_line()+
  labs(x = 'Dag van het Jaar',
       y = 'Bowen ratio',
       title = 'Bowen Ratio voor 2008')

```

### Step 4: Carbon fluxes
Carbon balance, carbon closure

- **Plot 6**: mean annual cycles for GPP, Reco and NEE, all in one plot
- Interpret the annual cycles. Is the site more a carbon source (NEE > 0) or a carbon sink (NEE < 0)?
```{r}
#PLOT 6
P6_data <- pivot_longer(averages_per_day_year, cols=c("NEE", "RECO", "GPP"),
                        names_to = 'variables',
                        values_to = 'values')
ggplot(P6_data, aes(x=DayOfYear, y=values, col=variables))+
  geom_line()+
  labs(x = 'Dag van het Jaar', y= 'C fluxen (kgC/m²/jaar)',
       title = 'gemiddelde C fluxen')+
  geom_hline(yintercept=0)+
  scale_color_discrete(name='variabelen')

```


### Step 5: Interpretation
Climate type, links between variables, relation with ecosystem 

- **Plot 7**: Are there links between different meteorological variables?
- **Plot 8**: Are some of the land fluxes (GPP, Reco, NEE) correlated with one or several meteorological variables?
- Compare climate type with that on other locations (e.g. other groups, textbook examples, papers, ...)
- Comment on extreme events, inter-annual variability, anomalies, …
- What is the relation between the climate and the site characteristics? What is the role of vegetation in this climate system?
- Any surprising details?
```{r}
#PLOT 7
P7_data <- select(extended_data, c('T (°C)', 'nacht T (°C)',
                                   'dag T (°C)','SW down (W/m²)',
                                   'LW down (W/m²)', 'VPD (hPa)',
                                   'regen (mm)', 'wind (m/s)',
                                   'bodem T (°C)'))
chart.Correlation(P7_data)

#PLOT 8
P8_data <- extended_data%>%
  filter(jaar>2006)%>%
  select(c('T (°C)', 'nacht T (°C)','dag T (°C)','SW down (W/m²)',
           'LW down (W/m²)', 'VPD (hPa)', 'regen (mm)', 
           'wind (m/s)','bodem T (°C)', "NEE (kgC/m²/jaar)",
           "GPP (kgC/m²/jaar)", "RECO (kgC/m²/jaar)"))
chart.Correlation(P8_data)

```


### Useful R cheatsheets
1. Summary of Rstudio functions: https://raw.githubusercontent.com/rstudio/cheatsheets/master/rstudio-ide.pdf
2. R basic commands: https://raw.githubusercontent.com/rstudio/cheatsheets/master/base-r.pdf
3. Data visualization: https://raw.githubusercontent.com/rstudio/cheatsheets/master/data-visualization-2.1.pdf
4. Work with dates: https://raw.githubusercontent.com/rstudio/cheatsheets/main/lubridate.pdf
5. More R cheatsheets: https://rstudio.com/resources/cheatsheets/

