---
title: "Flux analysis"
date: "10/11/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r libraries}
library(dplyr)
library(tidyr)
library(ggplot2)
library(naniar)
```

**Meteorological data:** 
(UFORA) You only need the data corresponding to your study site.

- Time of the measurement (YEAR, MONTH, and DAY)
- Daily average temperature T (°C)
- Daily average night temperature, T_NIGHT (°C)
- Daily average day temperature, T_DAY (°C)
- Incoming shortwave radiation, SWDOWN (W/m²)
- Incoming longwave radiation, LWDOWN (W/m²)
- Vapor pressure deficit, VPD (hPa)
- Rainfall, RAIN (mm)
- Windspeed, WIND (m/s)
- Net radiation, RN (W/m²)
- Soil temperature, TSOIL (°C)
- Soil heat flux, GF (W/m²)
- Latent heat flux QLE (W/m²)
- Sensible heat flux QH (W/m²)
- Net Ecosystem Exchange, NEE (gC/m²/d)
- Gross primary production, GPP (gC/m²/d)
- Ecosystem respiration, Reco (gC/m²/d)

**Additional helpful Metadata:**

- World map of Köppen-Geiger climate classification (PDF)
- Site information: Find from [fluxnet sites list](https://fluxnet.org/sites/site-list-and-pages) or just google the name of your site


### Step 1: Explore your dataset 
Data cleanup, units conversion; typical ranges, annual sums and averages  

- Check and filter missing (NA) or impossible (e.g. very larges, -9999) values
- Convert variables into easy interpretable units (land fluxes into kgC /m²/year or tC /ha/year)
- Calculate the minimal and maximal value for each variable, as a quick check
- Calculate annual sum of precipitation, and the annual mean for each other variable. 

```{r Loading your data}
# Loading data
Fluxname <- "ZM-Mon_daily"
data.file <- file.path("./",paste0(Fluxname,".csv"))
data <- read.csv(data.file,header = TRUE,sep = ";")

```

```{r cleaning data}
# cleaning the data
data <- data%>%
  replace_with_na_if(.predicate = is.double,condition = ~.x < -1000)


```

```{r Unit conversion}
# Editing the units
data <- data %>%
  mutate(
    NEE = NEE*10^-3*365,
    RECO = RECO*10^-3*365,
    GPP = GPP*10^-3*365
  )
min <- data.frame(as.list(apply(data, 2, min, na.rm=TRUE)))
max <- data.frame(as.list(apply(data, 2, max, na.rm=TRUE)))
summarie <- rbind(min, max)

precip_cumsum <- data %>%
  group_by(YEAR)%>%
  summarise(sum=sum(RAIN))
```


### Step 2: Mean annual cycles
Seasonal cycle, patterns, extreme events 

- **Plot 1**: Produce mean seasonal cycle graphs (= you compile your data to obtain one average year) for each meteorological variable (in a single plot)
- Check the correspondence of these plots: do patterns concur or differ? Why? Try to explain the causal relationship. 
- **Plot 2**: Compare your average year with every single year of data. Are there any extreme events in your data? Document it.
- **Plot 3**: Using bar plots, plot the average precipitation yearly cycle
```{r }


```


### Step 3: Surface energy balance
Radiation balance, energy closure, Bowen ratio 

- **Plot 4**: Plot mean annual cycles for Rn, Qh, Qle, GF (W/m²), in one graph  and interpret the annual cycles.
- **Plot 5**: Scatterplot of energy balance (Qh+Qle vs. Rn-GF) (W/m²), select only 1 year of your dataset for this plot, fit a linear regression line (include equation + R²). Compare with 1:1 line.
- Analyze the closure of the energy balance. How much is the energy gap? What causes this? Did you expect such magnitude? Refer to literature in your answer.
- Calculate the Bowen ratio (= Qh/Qle).
```{r }

```

### Step 4: Carbon fluxes
Carbon balance, carbon closure

- **Plot 6**: mean annual cycles for GPP, Reco and NEE, all in one plot
- Interpret the annual cycles. Is the site more a carbon source (NEE > 0) or a carbon sink (NEE < 0)?
```{r}


```


### Step 5: Interpretation
Climate type, links between variables, relation with ecosystem 

- **Plot 7**: Are there links between different meteorological variables?
- **Plot 8**: Are some of the land fluxes (GPP, Reco, NEE) correlated with one or several meteorological variables?
- Compare climate type with that on other locations (e.g. other groups, textbook examples, papers, ...)
- Comment on extreme events, inter-annual variability, anomalies, …
- What is the relation between the climate and the site characteristics? What is the role of vegetation in this climate system?
- Any surprising details?
```{r}


```


### Useful R cheatsheets
1. Summary of Rstudio functions: https://raw.githubusercontent.com/rstudio/cheatsheets/master/rstudio-ide.pdf
2. R basic commands: https://raw.githubusercontent.com/rstudio/cheatsheets/master/base-r.pdf
3. Data visualization: https://raw.githubusercontent.com/rstudio/cheatsheets/master/data-visualization-2.1.pdf
4. Work with dates: https://raw.githubusercontent.com/rstudio/cheatsheets/main/lubridate.pdf
5. More R cheatsheets: https://rstudio.com/resources/cheatsheets/

